Critical assessment of Linear-Response Excitations in Molecular system\\ or \\ On the Localization of the Linear-Response excitations in Open Systems

### 2018-01-29 LRDraft.tex
update

% notation for the operators
\newcommand{\op}[1]{\hat {#1}}
\newcommand{\sop}[1]{\op{\op {#1}}}
\newcommand{\commutator}[2]{\left[ {#1} , {#2} \right]}
\newcommand{\trace}[1]{\mathrm{tr}\left(#1\right)}
\newcommand{\argument}[1]{\ifthenelse{\isempty{#1}{}}{}{(#1)}}
\newcommand{\matop}[1]{\mathbf{#1}}
\newcommand{\optr}[1]{\check #1}

@@

The formal scalar products that express the linear response functional \eqref{LinearResponseFunctDef1} can be evaluated in the same basis introduced above, that is: 
%For instance the first addend of \eqref{LinearResponseFunctDef2} reads: 
%\be\lb{LinearResponseFunctDef3}
%\bra{f_p(-\omega)}\op O\ket{\psi_p} = \sum_\mu \brket{f_p(-\omega)}{\mu}\bra{\mu}\op O\ket{\psi_p} \;.
%\ee
\be\lb{LinearResponseFunctDef3}
{\cal R}_{\op O}(\omega) = 
\sum_p\sum_\mu\left(\brket{f_p(-\omega)}{\mu}\bra{\mu}\op O\ket{\psi_p} + \bra{\psi_p}\op O\ket{\mu}\brket{\mu}{f_p(\omega)}\right) \;.
\ee
The task of the linear response formalism to provide reliable results for the expectation value of $\op O$ can be reformulated in the problem of finding an accurate and unbiased
expression for the numerical coefficients appearing in \eqref{LinearResponseFunctDef3}. The features of the computational basis described above ensures that this requirement should
be satisfied for $\bra{\mu}\op O\ket{\psi_p}$ (and its hermitian conjugates). 
On the contrary, the problem of the reliable reconstruction of the fluctuation states through the sampling of the coefficients $\brket{f_p(-\omega)}{\mu}$ and $\brket{\mu}{f_p(\omega)}$
turns out to be highly non-trivial, since the behavior of the fluctuation states strongly depends on $\omega$, as will be readily clarified in the next section.

%On the contrary, the most interesting point of this discussion concerns with the assessment of the 
%reliable sampling of the fluctuation states through the coefficients $\brket{f_p(-\omega)}{\mu}$ and $\brket{\mu}{f_p(\omega)}$. The problem turns out to be highly is non-trivial 
%since the behavior of the fluctuation states strongly depends on $\omega$, as will be readily clarified in the next section.

%$\brket{f_p(-\omega)}{\mu}$ and $\brket{\mu}{f_p(\omega)}$ in the domain in which the matrix elements $\bra{\mu}\op O\ket{\psi_p}$ (and its hermitian 
%conjugates) are not vanishing. 
%The task of the linear response formalism to provide reliable results for the expectation value of $\op O$ can be reformulated in the problem of finding an accurate and unbiased
%expression for the coefficients $\brket{f_p(-\omega)}{\mu}$ and $\brket{\mu}{f_p(\omega)}$ in the domain in which the matrix elements $\bra{\mu}\op O\ket{\psi_p}$ (and its hermitian 
%conjugates) are not vanishing.

% We assume that the ground state problem has been already solved and that the occupied orbitals  introduce a computational basis $\{\ket{\mu}\}$, composed by a \emph{finite}
% set of functions
% 
% Here we are dealing with a linear response problem and

% The quality of the basis can be evaluated by the requirement that the numerical coefficients appearing in the r.h.s of \eqref{LinearResponseFunctDef3} provide a reliable reconstruction 
% of the formal scalar product.

% Here we are assuming that the ground state properties of the unperturbed system have been computed in the same basis introduced above. This means that the 
% feature of the basis, that for instance for real space basis are given by the grid spacing and number of points (that determines the dimension simulation box), are chosen so that the 
% coefficients $\brket{\mu}{\psi_p}$ represent a valid discretization of the occupied orbitals. The same can be stated also for the matrix elements $\bra{\mu}\op O\ket{\psi_p}$, by assuming 
% that $\op O$ acts as a local operator on the occupied orbitals.

% aggiungere un preambolo a quello che segue. Soluzione usando le equazioni di Sternheimer e approccio con le eccitazioni....

\subsection{Sternheimer equations and localization properties of the fluctuation states}
%\subsection{Localization argument for the fluctuation states}
\lb{LocalizationFluctuationStates}

@@

\subsubsection{Criterion for the choice of the computational basis} %underlying

@@

%\subsection{Excitation operator and the analytic structure of the linear response}
\subsection{Linear response \emph{via} the construction of the excitation operators}

@@

%\subsubsection{Definition of the basis of transitions}
\subsubsection{Explicit sampling of the empty-states subspace}

@@

% We now introduce the notion of \emph{basis of transitions} and identify the operators than can be expanded in such a basis. 
% 
% We split the complete
% set of eigenstates of $\hnot$ in two sets. The first one is composed by the eigenvectors of $\hnot$ in the subspace of the occupied orbitals. The second 
% one is composed by the remaining eigenvectors, both in the bound and in the continuum region, and its elements are denoted as $\ket{\psi(\alpha)}$, where $\alpha$ 
% is a multi-index in the sense described in \ref{BasicDef}. The elements of these sets satisfy the following orthonormality relation: 
% \be\lb{orthonormalityDef1}
% \brket{\psi(\alpha)}{\psi_p} = \brket{\psi_p}{\psi(\alpha)} = 0 \;, \qq \brket{\psi_p}{\psi_q} = \delta_{pq} \;, \qq
% \brket{\psi(\alpha)}{\psi(\beta)} = \delta(\alpha-\beta) \;.
% \ee
% Obviously, the direct sum of these two sets coincides with the complete set of eigenstates of $\hnot$ and thus 
% constitute a complete basis.

@@

% \vspace{2cm}
% A generic operator can be decomposed in this basis and its formal expansion contains four distinct contribution of the form $\ketbra{\psi_p}{\psi_q}$, 
% $\ketbra{\psi(\alpha)}{\psi(\beta)}$, $\ketbra{\psi(\alpha)}{\psi_p}$, $\ketbra{\psi_p}{\psi(\alpha)}$. However an inspection of equations \eqref{rhoPrimeOmegaDef1}
% and \eqref{LiouvillianRhopomegaDef3} evidences that the relevant operators in the contest of the linear response are expressed as a commutator with $\dmnot$. 
% For this class of operators the only non vanishing contribution comes from the \emph{transition operators}:
% \be
% \excite{p}{\alpha} \equiv \ketbra{\psi(\alpha)}{\psi_p}\;, \qq
% \decay{\alpha}{p} \equiv \ketbra{\psi_p}{\psi(\alpha)} \;.
% \ee
% The orthonormality conditions \eqref{psiAlphaDef1} imply that:
% \be\lb{transitionOpOrthonormalityDef1}
% \trace{\excite{p}{\alpha}\excite{q}{\beta}} = \trace{\decay{\alpha}{p}\decay{\beta}{q}} = 0 \;, \qq
% \trace{\excite{p}{\alpha}\decay{\beta}{q}} = \delta_{pq}\delta(\alpha-\beta) \;,
% \ee
% so that the expansion of a suitable operator $\op O$ in the basis of transitions reads:
% \be\lb{OperatorTransBasisExpDef1}
% \op O = \sum_p\sint \dd \alpha\left(
% \excite{p}{\alpha}\trace{\decay{\alpha}{p}\op O} + \decay{\alpha}{p}\trace{\excite{p}{\alpha}\op O}
% \right) \;.
% \ee

@@

%\subsubsection{Spectral decomposition of the response density in the transitions basis}
%The Liouville equation \eqref{LiouvillianRhopomegaDef3} can be consistently written in the transition basis. 
% Indeed, its r.h.s is explicitly written
% as a commutator with $\dmnot$, while an inspection of the formal definition of the response density and of the Liouville superoperator evidences that
% they both contain a commutator with $\dmnot$. Written in the basis of transitions Liouville equation becomes a linear matrix equation for the rank-2 
% vectors of the coefficients \eqref{OperatorTransitionCoeffDef1} of the operators in its left and right side:

@@

@@

\mat{ \trace{\decay{\alpha}{p}\commutator{\op\Phi(\omega)}{\dmnot}} \\ \trace{\excite{p}{\alpha}\commutator{\op\Phi(\omega)}{\dmnot}}}
\;.\nn

@@

The second addend in the l.h.s can be further analyzed by writing the transition basis expansion of $\dm'$ inside the traces. Performing this procedure
we can recast this quantity as the action of a 2-dimensional matrix on the rank-2 vector of the response density. So we have: 

@@

\sket{p\alpha}{\commutator{\op\Phi(\omega)}{\dmnot}} \;.

@@

\subsubsection{Spectral decomposition of the response density in the transitions basis}

This formulation of the Liouville equation is the starting point for the derivation of the spectral decomposition of the response density operator
in terms of the eigenvectors if the Liouville operator. 
The first step in the achievement of this result is provided by the construction of the eigenvectors of the matrix \eqref{LiouvilleMatrixTransBasisDef1}.
Following the notation previously defined we introduce the rank-2 vectors $\sket{p\alpha}{E}$ and $\sbra{p\alpha}{\tilde E}$ that represent, by definition,
the right and left eigenvectors $\optr{L}$, respectively, so that:
\be\lb{LiouvMatrixEigenvectorDef1}
\sum_q \sint \dd \beta \optr{L}^{p \alpha}_{q\beta} \sket{q \beta}{E} = E \sket{p\alpha}{E} \;, \quad
\sum_p \sint \dd \alpha \sbra{p\alpha}{\tilde E} \optr{L}^{p \alpha}_{q\beta} = E \sbra{q\beta}{\tilde E}\;.
\ee
In general, left and right eigenvectors do not coincide, but it is always possible to choose they normalization so that a orthonormality conditions among the
elements of left and right spaces holds, namely:
\be\lb{ExcitationOrthonormalityDef1}
\sum_p \sint \dd \alpha \sbra{p\alpha}{\tilde E}\sket{p\alpha}{E'} \equiv \dbraket{\tilde E}{E'} = \delta(E-E') \;.
\ee
Once that the eigenvectors are known, the spectral decomposition of the resolvent of $\optr{L}$ can be provided:  
\be
\left[\left(\bbomega-\optr{L}\right)^{-1}\right]^{p\alpha}_{q\beta} =\sint\dd E
\frac{\sket{p\alpha}{E}\sbra{q\beta}{\tilde E}}{\bbomega-E} \;,
\ee
so equation \eqref{LiouvilleTransBasisDef1} can be inverted and we obtain the response density written in basis of transistion, in the form:
% \be\lb{RhopExcitationTransBasisDef1}
% \sket{p\alpha}{\dm'(\omega)} = \sint\dd E \sket{p\alpha}{E}
% \frac{\sum_q \sint \dd \beta\sbra{q\beta}{\tilde E}
% \sket{q\beta}{\commutator{\op\Phi(\omega)}{\dmnot}}}{\bbomega-E} \;,
% \ee
% that can be compactly rewritten by using the index-less notation:
\be\lb{RhopExcitationTransBasisDef1}
\sket{p\alpha}{\dm'(\omega)} = 
\sint\dd E \sket{p\alpha}{E}
\frac{\dbraket{\tilde E}{\commutator{\op\Phi(\omega)}{\dmnot}}}{\bbomega-E} \;.
\ee
% \be\lb{RhopExcitationTransBasisDef2}
% \dket{\dm'(\omega)} = 
% \sint\dd E \dket{E}
% \frac{\dbraket{\tilde E}{\commutator{\op\Phi(\omega)}{\dmnot}}}{\bbomega-E} \;.
% \ee

\subsubsection{Definition of excitation operators}

We introduce the explicit representation of the eigenvectors op $\optr{L}$:
\be\lb{XYcoeffDef1}
\sket{p\alpha}{E} \equiv \mat{X_E^{p\alpha} \\ Y_E^{p\alpha}} \;, \qq
\sbra{p\alpha}{\tilde E} \equiv \mat{\tilde X_{E\,p\alpha} & \tilde Y_{E\,p\alpha}} \;,
\ee
and on the basis of these elements we define the \emph{excitation operators}:
\begin{align}\lb{ExctitationOperatorsDef1}
\op E &\equiv \dbraket{\cdot}{E} = 
\sum_p \sint \dd \alpha \left(\excite{p}{\alpha} X_E^{p\alpha} + \decay{\alpha}{p} Y_E^{p\alpha}\right) 
\;,\nn \\
\op{\tilde E} &\equiv \dbraket{\tilde E}{\cdot} = \sum_p \sint \dd \alpha \left(\decay{\alpha}{p} \tilde X_{E\,p\alpha} + 
\excite{p}{\alpha}\tilde Y_{E\,p\alpha}  \right)\;.
\end{align}
The orthonormality condition \eqref{ExcitationOrthonormalityDef1} implies that:
\be
\trace{\op{\tilde E} \op E'} = \sum_p\sint \dd\alpha \left(\tilde X_E^{p\alpha} X_{E'\,p\alpha} +
\tilde Y_E^{p\alpha} Y_{E'\,p\alpha}\right) = \delta(E-E') \;. 
\ee
Excitation operators are the right and left eigenvectors of the Liouville superoperator, as directly follows from their
definition. Indeed for $\op E$ we have:
\begin{align}\lb{opEeigenvalueDef1}
\Liouv \op E & = \sum_p \sint \dd \alpha \left(\trace{\decay{\alpha}{p} \Liouv \op E} \excite{p}{\alpha} + 
\decay{\alpha}{p} \trace{ \excite{p}{\alpha} \Liouv \op E}  \right) = \nn \\
&= \sum_p \sint  \dd \alpha \sbra{p\alpha}{\cdot} \sum_q \sint  \dd \beta \optr{L}^{p \alpha}_{q\beta} \sket{q\beta}{E} 
= E \sum_p \sint \dd \alpha \sbra{p\alpha}{\cdot} \sket{p\alpha}{E} = E \, \op E \;,	
\end{align}
whereas the left action on $\op{\tilde E}$ provides:
\begin{align}\lb{optildeEeigenvalueDef1}
 \op{\tilde E} \Liouv &=  \sum_p \sint \dd \alpha \left(
\trace{\op{\tilde E}\Liouv\excite{p}{\alpha}}\decay{\alpha}{p} +  
\trace{\op{\tilde E}\sop{\mathcal L}\decay{\alpha}{p}} \excite{p}{\alpha} \right) = 
\nn \\ &= \sum_p \sint  \dd \alpha  \sum_q \sint  \dd \beta
\sbra{q\beta}{\tilde E}\optr{L}^{q\beta}_{p\alpha}\sket{p\alpha}{\cdot} = 
E\sum_p \sint \dd \alpha \sbra{p\alpha}{\tilde E} \sket{p\alpha}{\cdot}  =
E \op{\tilde E} \;.
\end{align}
The introduction of the excitation operators allows to provide a complete operator version of equation \eqref{RhopExcitationTransBasisDef1}, that is:
\be\lb{RhopExcitationDef1}
\dm'(\omega) = 
\sint\dd E \op E
\frac{\trace{\op{\tilde E}\commutator{\op\Phi(\omega)}{\dmnot}}}{\bbomega-E} \;,
\ee
where the algebraic equivalence between the trace written in \eqref{RhopExcitationDef1} and the numerator of \eqref{RhopExcitationTransBasisDef1}
can be easily proved.

@@

We introduce the \emph{excited states} defined via the linear combination of the basis set orthogonal to $\dmnot$ with the
coefficients that appears in \eqref{XYcoeffDef1}, that is:
\begin{align}\lb{ExcitonicStatesDef1}
\ket{\phi^E_p} &\equiv \sint \dd \alpha X_{E}^{p \alpha} \ket{\psi(\alpha)} &
\tbra{\phi^E_p} &\equiv \sint \dd \alpha {\tilde X_{E}^{p \alpha}} \bra{\psi(\alpha)} \;, \nn\\
\bra{\chi^E_p} &\equiv \sint \dd \alpha  Y_{E}^{p \alpha} \bra{\psi(\alpha)} &
\tket{\chi^E_p} &\equiv \sint \dd \alpha {\tilde Y}_{E}^{p\alpha} \ket{\psi(\alpha)} \;.
\end{align}
Thanks to these definitions we can express the excitation operators as follows:
\be\lb{ExctitationOperatorsDef2}
\op E = \sum_p \left( \ket{\phi^E_p}\bra{\psi_p} + \ket{\psi_p} \bra{\chi^E_p}\right)\;, \qq
\op {\tilde E} = \sum_p \left( \ket{\psi_p} \tbra{\phi^E_p} + \tket{\chi^E_p}\bra{\psi_p} \right)\;.
\ee
Starting from these equation we can derive the excitation expansion of the fluctuation states. Indeed an inspection of formula \eqref{rhoPrimeFluctuationStateDef1}
show that fluctuation states can be extracted from the response density operator as follows:
\be
\ket{f_p(\omega)} = \dm'(\omega)\ket{\psi_p} \;, \qq
\bra{f_p(-\omega)} = \bra{\psi_p}\dm'(\omega) \;, \nn
\ee
so plugging equations \eqref{ExctitationOperatorsDef2} in the formula \eqref{RhopExcitationDef1} and performing these operations we obtain:
\be\lb{FluctStateExcitationDef1}
\ket{f_p(\omega)} = \sint\dd E \frac{\trace{\op{\tilde E}\commutator{\op\Phi(\omega)}{\dmnot}}}{\bbomega-E} \ket{\phi^E_p} \;, \qq
\bra{f_p(-\omega)} = \sint\dd E \bra{\chi^E_p}\frac{\trace{\op{\tilde E}\commutator{\op\Phi(\omega)}{\dmnot}}}{\bbomega-E} 
\ee

@@

@@

% \section{Introduction}
% 
% \subsection{How to add Comments}
% 
% Comments can be added to your project by clicking on the comment icon in the toolbar above. % * <john.hammersley@gmail.com> 2016-07-03T09:54:16.211Z:
% %
% % Here's an example comment!
% %
% To reply to a comment, simply click the reply button in the lower right corner of the comment, and you can close them when you're done.
% 
% Comments can also be added to the margins of the compiled PDF using the todo command\todo{Here's a comment in the margin!}, as shown in the example on the right. You can also add inline comments:
% 
% \todo[inline, color=green!40]{This is an inline comment.}
% 
% \subsection{How to add Tables}
% 
% Use the table and tabular commands for basic tables --- see Table~\ref{tab:widgets}, for example. 
% 
% \begin{table}
% \centering
% \begin{tabular}{l|r}
% Item & Quantity \\\hline
% Widgets & 42 \\
% Gadgets & 13
% \end{tabular}
% \caption{\label{tab:widgets}An example table.}
% \end{table}
% 
% \subsection{How to add Citations and a References List}
% 
% You can upload a \verb|.bib| file containing your BibTeX entries, created with JabRef; or import your \href{https://www.overleaf.com/blog/184}{Mendeley}, CiteULike or Zotero library as a \verb|.bib| file. You can then cite entries from it, like this: \cite{greenwade93}. Just remember to specify a bibliography style, as well as the filename of the \verb|.bib|.
% 
% You can find a \href{https://www.overleaf.com/help/97-how-to-include-a-bibliography-using-bibtex}{video tutorial here} to learn more about BibTeX.
% 
% We hope you find Overleaf useful, and please let us know if you have any feedback using the help menu above --- or use the contact form at \url{https://www.overleaf.com/contact}!

>>>>

% notation for the operators
\newcommand{\op}[1]{\hat {#1}}
\newcommand{\sop}[1]{\op{\op {#1}}}
\newcommand{\commutator}[2]{\left[ {#1} , {#2} \right]}
\newcommand{\trace}[1]{\mathrm{tr}\left(#1\right)}
\newcommand{\argument}[1]{\ifthenelse{\isempty{#1}{}}{}{(#1)}}
\newcommand{\matop}[1]{\mathbf{#1}}
\newcommand{\optr}[1]{\check #1}
\newcommand{\opskew}[1]{{\op {#1}}_{\perp}}

@@

The formal scalar products that express the linear response functional \eqref{LinearResponseFunctDef1} can be evaluated in the same basis introduced above, that is: 
\be\lb{LinearResponseFunctDef3}
{\cal R}_{\op O}(\omega) = 
\sum_p\sum_\mu\left(\brket{f_p(-\omega)}{\mu}\bra{\mu}\op O\ket{\psi_p} + \bra{\psi_p}\op O\ket{\mu}\brket{\mu}{f_p(\omega)}\right) \;.
\ee
The task of the linear response formalism to provide reliable results for the expectation value of $\op O$ can be reformulated in the problem of finding an accurate and unbiased
expression for the numerical coefficients appearing in \eqref{LinearResponseFunctDef3}. The features of the computational basis described above ensures that this requirement should
be satisfied for $\bra{\mu}\op O\ket{\psi_p}$ (and its hermitian conjugates). 
On the contrary, the problem of the reliable reconstruction of the fluctuation states through the sampling of the coefficients $\brket{f_p(-\omega)}{\mu}$ and $\brket{\mu}{f_p(\omega)}$
turns out to be highly non-trivial, since the behavior of the fluctuation states strongly depends on $\omega$, as will be readily clarified in the next section.

\subsection{Sternheimer equations and localization properties of the fluctuation states}
\lb{LocalizationFluctuationStates}

@@

\subsubsection{Criterion for the choice of the computational basis}

@@

\subsection{Linear response \emph{via} the construction of the excitation operators}

@@

\subsubsection{Explicit sampling of the empty-states subspace}

@@

@@

@@

@@

\be\lb{LiouvillianRhopomegaDef4}

@@

\mat{ \trace{\decay{\alpha}{p}\opskew{\Phi}(\omega)} \\ \trace{\excite{p}{\alpha}\opskew{\Phi}(\omega)}}
\;,\nn

@@

where we have introduced a compact notation to deal with the commutator of on operator with the unperturbed density: 
\opskew{O} = \commutator{\op O}{\dmnot}\;.
The second addend in the l.h.s of \eqref{LiouvillianRhopomegaDef4} can be further analyzed by writing the transition basis expansion of $\dm'$ inside the traces. 
Performing this procedure we can recast this quantity as the action of a 2-dimensional matrix on the rank-2 vector of the response density. So we have: 

@@

\sket{p\alpha}{\opskew{\Phi}(\omega)} \;.

@@

\subsubsection{Spectral decomposition of the response density in the transitions basis. Definition of excitation operators}

The formulation of the Liouville equation in the basis of transitions is the starting point for the derivation of the spectral decomposition of the response density operator
in terms of the eigenvectors if the Liouville operator. The first step in the achievement of this result is provided by the construction of the eigenvectors of the matrix 
\eqref{LiouvilleMatrixTransBasisDef1}. Following the notation previously defined we introduce the rank-2 vectors $\sket{p\alpha}{E}$ and $\sbra{p\alpha}{\tilde E}$ that represent, 
by definition, the right and left eigenvectors $\optr{L}$, respectively, so that:
\be\lb{LiouvMatrixEigenvectorDef1}
\sum_q \sint \dd \beta \optr{L}^{p \alpha}_{q\beta} \sket{q \beta}{E} = E \sket{p\alpha}{E} \;, \quad
\sum_p \sint \dd \alpha \sbra{p\alpha}{\tilde E} \optr{L}^{p \alpha}_{q\beta} = E \sbra{q\beta}{\tilde E}\;.
In general, left and right eigenvectors do not coincide, but it is always possible to choose they normalization so that a orthonormality conditions among the
elements of left and right spaces holds, namely:
\be\lb{ExcitationOrthonormalityDef1}
\sum_p \sint \dd \alpha \sbra{p\alpha}{\tilde E}\sket{p\alpha}{E'} \equiv \dbraket{\tilde E}{E'} = \delta(E-E') \;.
Once that the eigenvectors are known, the spectral decomposition of the resolvent of $\optr{L}$ can be provided:  
\left[\left(\bbomega-\optr{L}\right)^{-1}\right]^{p\alpha}_{q\beta} =\sint\dd E
\frac{\sket{p\alpha}{E}\sbra{q\beta}{\tilde E}}{\bbomega-E} \;,
so equation \eqref{LiouvilleTransBasisDef1} can be inverted and we obtain the response density written in basis of transition, in the form:
\be\lb{RhopExcitationTransBasisDef1}
\sket{p\alpha}{\dm'(\omega)} = 
\sint\dd E \sket{p\alpha}{E}
\frac{\dbraket{\tilde E}{\opskew{\Phi}(\omega)}}{\bbomega-E} \;.
%\subsubsection{Definition of excitation operators}
It is possible to express formula \eqref{RhopExcitationTransBasisDef1} as a true operator equation. First of all we show that the rank-2 vectors $\sket{p\alpha}{E}$ 
and $\sbra{p\alpha}{\tilde E}$ can be interpreted as the transition basis coefficients of the right and left \emph{eigenstates} of the Liouville superoperator.   
In order to do so we introduce the explicit representation of the eigenvectors op $\optr{L}$:
\be\lb{XYcoeffDef1}
\sket{p\alpha}{E} \equiv \mat{X_E^{p\alpha} \\ Y_E^{p\alpha}} \;, \qq
\sbra{p\alpha}{\tilde E} \equiv \mat{\tilde X_{E\,p\alpha} & \tilde Y_{E\,p\alpha}} \;,
and, on the basis of these elements, we define the \emph{excitation operators}:
\begin{align}\lb{ExctitationOperatorsDef1}
\op E &\equiv \dbraket{\cdot}{E} = 
\sum_p \sint \dd \alpha \left(\excite{p}{\alpha} X_E^{p\alpha} + \decay{\alpha}{p} Y_E^{p\alpha}\right) 
\;,\nn \\
\op{\tilde E} &\equiv \dbraket{\tilde E}{\cdot} = \sum_p \sint \dd \alpha \left(\decay{\alpha}{p} \tilde X_{E\,p\alpha} + 
\excite{p}{\alpha}\tilde Y_{E\,p\alpha}  \right)\;.
\end{align}
The orthonormality condition \eqref{ExcitationOrthonormalityDef1} implies that:
\trace{\op{\tilde E} \op E'} = \sum_p\sint \dd\alpha \left(\tilde X_E^{p\alpha} X_{E'\,p\alpha} +
\tilde Y_E^{p\alpha} Y_{E'\,p\alpha}\right) = \delta(E-E') \;. 
Excitation operators are the right and left eigenvectors of the Liouville superoperator, as directly follows from their
definition. Indeed for $\op E$ we have:
\begin{align}\lb{opEeigenvalueDef1}
\Liouv \op E & = \sum_p \sint \dd \alpha \left(\trace{\decay{\alpha}{p} \Liouv \op E} \excite{p}{\alpha} + 
\decay{\alpha}{p} \trace{ \excite{p}{\alpha} \Liouv \op E}  \right) = \nn \\
&= \sum_p \sint  \dd \alpha \sbra{p\alpha}{\cdot} \sum_q \sint  \dd \beta \optr{L}^{p \alpha}_{q\beta} \sket{q\beta}{E} 
= E \sum_p \sint \dd \alpha \sbra{p\alpha}{\cdot} \sket{p\alpha}{E} = E \, \op E \;,	
\end{align}
whereas the left action on $\op{\tilde E}$ provides:
\begin{align}\lb{optildeEeigenvalueDef1}
 \op{\tilde E} \Liouv &=  \sum_p \sint \dd \alpha \left(
\trace{\op{\tilde E}\Liouv\excite{p}{\alpha}}\decay{\alpha}{p} +  
\trace{\op{\tilde E}\sop{\mathcal L}\decay{\alpha}{p}} \excite{p}{\alpha} \right) = 
\nn \\ &= \sum_p \sint  \dd \alpha  \sum_q \sint  \dd \beta
\sbra{q\beta}{\tilde E}\optr{L}^{q\beta}_{p\alpha}\sket{p\alpha}{\cdot} = 
E\sum_p \sint \dd \alpha \sbra{p\alpha}{\tilde E} \sket{p\alpha}{\cdot}  =
E \op{\tilde E} \;.
\end{align}
The introduction of the excitation operators allows us to provide a complete operator version of equation \eqref{RhopExcitationTransBasisDef1}, that is:
\be\lb{RhopExcitationDef1}
\dm'(\omega) = \dbraket{\cdot}{\dm'(\omega)} =
\sint\dd E \op E
\frac{\trace{\op{\tilde E}\opskew{\Phi}(\omega)}}{\bbomega-E} \;,
where the algebraic equivalence between the trace written in \eqref{RhopExcitationDef1} and the numerator of \eqref{RhopExcitationTransBasisDef1}
can be easily proved.

Thanks to the usage of the basis of transition, we have been able to express the response density as a linear superposition of excitations $\op E$. The perturbing field 
determines the relative contribution of each excitation through the weighting coefficients in the numerator of \eqref{RhopExcitationDef1}.
The fundamental question raised in the present paper, namely the ....reliable assessment of the results of the linear response can thus be reformulated in this language.

@@

We introduce the \emph{excited states} defined via the linear combination of the basis set orthogonal to $\dmnot$ with the
coefficients that appears in \eqref{XYcoeffDef1}, that is:
\begin{align}\lb{ExcitonicStatesDef1}
\ket{\phi^E_p} &\equiv \sint \dd \alpha X_{E}^{p \alpha} \ket{\psi(\alpha)} &
\tbra{\phi^E_p} &\equiv \sint \dd \alpha {\tilde X_{E}^{p \alpha}} \bra{\psi(\alpha)} \;, \nn\\
\bra{\chi^E_p} &\equiv \sint \dd \alpha  Y_{E}^{p \alpha} \bra{\psi(\alpha)} &
\tket{\chi^E_p} &\equiv \sint \dd \alpha {\tilde Y}_{E}^{p\alpha} \ket{\psi(\alpha)} \;.
\end{align}
Thanks to these definitions we can express the excitation operators as follows:
\be\lb{ExctitationOperatorsDef2}
\op E = \sum_p \left( \ket{\phi^E_p}\bra{\psi_p} + \ket{\psi_p} \bra{\chi^E_p}\right)\;, \qq
\op {\tilde E} = \sum_p \left( \ket{\psi_p} \tbra{\phi^E_p} + \tket{\chi^E_p}\bra{\psi_p} \right)\;.
\ee
Starting from these equation we can derive the excitation expansion of the fluctuation states. Indeed an inspection of formula \eqref{rhoPrimeFluctuationStateDef1}
show that fluctuation states can be extracted from the response density operator as follows:
\be
\ket{f_p(\omega)} = \dm'(\omega)\ket{\psi_p} \;, \qq
\bra{f_p(-\omega)} = \bra{\psi_p}\dm'(\omega) \;, \nn
\ee
so plugging equations \eqref{ExctitationOperatorsDef2} in the formula \eqref{RhopExcitationDef1} and performing these operations we obtain:
\be\lb{FluctStateExcitationDef1}
\ket{f_p(\omega)} = \sint\dd E \frac{\trace{\op{\tilde E}\opskew{\Phi}(\omega)}}{\bbomega-E} \ket{\phi^E_p} \;, \qq
\bra{f_p(-\omega)} = \sint\dd E \bra{\chi^E_p}\frac{\trace{\op{\tilde E}\opskew{\Phi}(\omega)}}{\bbomega-E} 
\ee

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@
